---
title: "NYC Schools Perceptions"
author: "Gavin Shin"
date: "2023-10-26"
output: html_document
---

In this project, we will answer to the following questions.

1. Do student, teacher, and parent perceptions of NYC school quality appear to be related to demographic and academic success metrics?

2. Do students, teachers, and parents have similar perceptions of NYC school quality?


Load the required packages for this analysis
```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(purrr)
library(tidyr)
library(stringr)
```

Import data into R

```{r}
survey1<-read_tsv("masterfile11_gened_final.txt")
survey2<-read_tsv("masterfile11_d75_final.txt")
combined <- read_csv("combined.csv")
```

Brief look of Data

```{r}
head(survey1,5)
head(survey2,5)
head(combined,5)
```
Include necessary columns for the analysis and combine those two 
```{r}
survey1_new <- survey1 %>%
                  filter(schooltype == "High School") %>%
                  select(dbn:aca_tot_11)
surevy2_new <- survey2 %>%
                  select(dbn:aca_tot_11)
combined_survey <- survey1_new %>%
                      bind_rows(surevy2_new)

```
Join combined_survey and combined dataframes
Use left join to keep all information from combined dataframe and corresponding information from survey
```{r}
combined_survey <- combined_survey %>%
                        rename(DBN = dbn)

combined_join <- combined %>%
                        left_join(combined_survey, by = "DBN")
```

Look for interesting relationships between variables "avg_sat_score" and convert it to tibble for efficient analysis
```{r}
cor_mat <- combined_join %>%
              select(avg_sat_score, saf_p_11:aca_tot_11) %>%
              cor(use = "pairwise.complete.obs")
cor_tib <- cor_mat %>%
              as_tibble(rownames = "variables")
```
Filter correlation to find strong relationship greater 0.3 or less than 0.3

```{r}
strong_cor <- cor_tib %>%
              select(variables, avg_sat_score) %>%
              filter(avg_sat_score > 0.25 | avg_sat_score < -0.25)
```

Make a function of scatter plot to look at the relationship visually.
```{r}
graph <- function(x,y){
  ggplot(combined_join, aes_string(x = x, y= y))+
    geom_point()
}

x = strong_cor$variables[2:5]
y = "avg_sat_score"

map2(x,y,graph)

```
All of those variables show a week positive relationship with the variable avg_sat_score

Reshape the data to see if there is a different outcome based on who answers the survey (parents, students, teacher)
```{r}
reshape_combined_join <- combined_join %>%
                            pivot_longer(cols = saf_p_11:aca_tot_11,
                                         names_to = "survey_questions",
                                         values_to = "score")
```
Create new variables response_type and metric
```{r}
reshape_combined_join <- reshape_combined_join %>%
                            mutate(response_type = str_sub(survey_questions, 5,6)) %>%
                            mutate(metric = str_sub(survey_questions, 1,3))

reshape_combined_join <- reshape_combined_join %>%
                            mutate(response_type = 
                                    ifelse(response_type == "p_","parent",
                                      ifelse(response_type == "s_","student",
                                        ifelse(response_type == "t_", "teacher",
                                          ifelse(response_type == "to","total", "NA")))))
```

Creating graphs based on reponse_type
```{r}
reshape_combined_join %>%
  filter(response_type != "total")%>%
  ggplot(aes(x = metric, y = score, fill = response_type)) + 
  geom_boxplot()
```

There are difference to some extent based on the type of questions, but generally speaking, parents' responses are the highest and teacher and students responses are following respectively.